---
title: "Solution to Homework 6"
author: "STATS/DATASCI 531, Winter 2022"
output:
  html_document:
    toc: no
bibliography: ../bib531.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/cell-numeric.csl
---

* Various difficulties were encountered with installing Xcode and Fortran compilers for Mac and R-tools for Windows. However, with persistence and some problem-solving on Piazza, these were all resolved.

* The following code is adapted from the notes for Simulation-Based Inference for Epidemiological Dynamics [@sismid].

## Question 6.1: Simulating an SIR process 

```{r read_data, include=FALSE}
library(tidyverse)
library(pomp)
set.seed(70982)
meas = read.csv(paste0("https://kingaa.github.io/sbied/stochsim/",
                       "Measles_Consett_1948.csv")) 

# The following chunk is written by Aaron King, Carles Bretó and
# Edward Ionides, and is adapted from the lecture file.

sir_step <- Csnippet("
double dN_SI = rbinom(S,1-exp(-Beta*I/N*dt));
double dN_IR = rbinom(I,1-exp(-mu_IR*dt));
S -= dN_SI;
I += dN_SI - dN_IR;
R += dN_IR;
H += dN_IR;
")

sir_init <- Csnippet("
S = nearbyint(eta*N);
I = 1;
R = nearbyint((1-eta)*N);
H = 0;
")

dmeas <- Csnippet("
lik = dbinom(reports,H,rho,give_log);
")

rmeas <- Csnippet("
reports = rbinom(H,rho);
")

measSIR = meas %>%
  select(week,reports=cases) %>%
  filter(week<=42) %>%
  pomp(
    times="week",t0=0,
    rprocess=euler(sir_step,delta.t=1/7),
    rinit=sir_init,
    rmeasure=rmeas,
    dmeasure=dmeas,
    accumvars="H",
    statenames=c("S","I","R","H"),
    paramnames=c("Beta","mu_IR","eta","rho","N")
  )

# End of the adapted chunk 

```

In the simulated outbreaks, the overall incidence is much too low, and the outbreak dies out after only a few weeks. To attempt to simulate data for which the observed data is a more plausible realization, we might try increasing the force of infection.

```{r initial_plot, echo=FALSE}
# From source [1]
measSIR %>%
  simulate(params=c(Beta=20,mu_IR=0.5,rho=0.5,eta=0.03,N=38000),
    nsim=20,format="data.frame",include.data=TRUE) %>%
  ggplot(aes(x=week,y=reports,group=.id,color=.id=="data"))+
  geom_line()+
  guides(color="none")
```

Taking it farther....

```{r, echo=FALSE}
measSIR %>%
  simulate(params=c(Beta=40,mu_IR=0.5,rho=0.5,eta=0.03,N=38000),
    nsim=20,format="data.frame",include.data=TRUE) %>%
  ggplot(aes(x=week,y=reports,group=.id,color=.id=="data"))+
  geom_line()+
  guides(color="none")
```

While this increases the overall incidence, the epidemic is now peaking too quickly. To counteract this, we might try reducing the recovery rate

```{r, echo=FALSE}
measSIR %>%
  simulate(params=c(Beta=40,mu_IR=0.2,rho=0.5,eta=0.03,N=38000),
    nsim=20,format="data.frame",include.data=TRUE) %>%
  ggplot(aes(x=week,y=reports,group=.id,color=.id=="data"))+
  geom_line()+
  guides(color="none")
```

Additionally, we might have a look at the effects of changing the initial susceptible fraction, η. It's possible to get something not too awful to contemplate by just manipulating $\eta$, in fact:

```{r, echo=FALSE}
measSIR %>%
  simulate(params=c(Beta=15,mu_IR=0.5,rho=0.5,eta=0.06,N=38000),
    nsim=20,format="data.frame",include.data=TRUE) %>%
  ggplot(aes(x=week,y=reports,group=.id,color=.id=="data"))+
  geom_line()+
  guides(color="none")
```

## Question 6.2: Simulating an SEIR process  

The existing code may be modified as follows:

```{r SEIR_building, include=FALSE}
seir_step <- Csnippet("
  double dN_SE = rbinom(S,1-exp(-Beta*I/N*dt));
  double dN_EI = rbinom(E,1-exp(-mu_EI*dt));
  double dN_IR = rbinom(I,1-exp(-mu_IR*dt));
  S -= dN_SE;
  E += dN_SE - dN_EI;
  I += dN_EI - dN_IR;
  R += dN_IR;
  H += dN_IR;
")


seir_init <- Csnippet("
  S = nearbyint(eta*N);
  E = 0;
  I = 1;
  R = nearbyint((1-eta)*N);
  H = 0;
")

measSIR %>%
  pomp(
    rprocess=euler(seir_step,delta.t=1/7),
    rinit=seir_init,
    paramnames=c("N","Beta","mu_EI","mu_IR","rho","eta"),
    statenames=c("S","E","I","R","H")
  ) -> measSEIR
```

Using the rough estimate that the latent period in measles is $8-10$ days, we take $\mu_{EI} \sim 0.8 wk^{-1}$ and $\mu_{IR} \sim 1.3 wk^{-1}$ (so as to have roughly the same generation time as before).

```{r mod1, echo=FALSE}
measSEIR %>%
  simulate(params=c(Beta=15,mu_EI=0.8,mu_IR=1.3,rho=0.5,eta=0.06,N=38000),
    nsim=20,format="data.frame",include.data=TRUE) %>%
  ggplot(aes(x=week,y=reports,group=.id,color=.id=="data"))+
  geom_line()+
  guides(color="none")
```

Again one can increase the force of infection:

```{r mod2, echo=FALSE}
measSEIR %>% 
  simulate(params=c(Beta=40,mu_EI=0.8,mu_IR=1.3,rho=0.5,eta=0.06,N=38000),
  nsim=20,format="data.frame",include.data=TRUE) %>%
  ggplot(aes(x=week,y=reports,group=.id,color=.id=="data"))+
  geom_line()+
  guides(color="none")
```

Now increase \(\beta\) to, by trial and error, 37. 

```{r mod3, echo=FALSE}
measSEIR %>%
simulate(
params=c(Beta=37,mu_EI=0.6,mu_IR=1,rho=0.7,eta=0.05,N=38000),
nsim=20,format="data.frame",include.data=TRUE
) -> sims
sims %>%
ggplot(aes(x=week,y=reports,group=.id,color=.id=="data"))+
geom_line()+
guides(color="none")
```

Notice that it seems that the simulated time series peak too late. We can decrease \(\eta\) back to 0.03, and increase \(\beta\) to 73.

```{r mod4, echo=FALSE}
measSEIR %>%
simulate(
params=c(Beta=73,mu_EI=0.6,mu_IR=1,rho=0.7,eta=0.03,N=38000),
nsim=20,format="data.frame",include.data=TRUE
) -> sims
sims %>%
ggplot(aes(x=week,y=reports,group=.id,color=.id=="data"))+
geom_line()+
guides(color="none")
```

Despite some uncertainties, this seems like a reasonable model for the time series. So we can choose \(\beta=73, \mu_{EI}=0.6,\mu_{IR}=1, \rho=0.7, \eta=0.03\) for the SEIR model.

----------------

### References

